---
title: "Ecosystem Description"
output:
  word_document: default
  html_document: default
always_allow_html: true
---


```{r setup, include=FALSE, results = "asis"}
library(tidyverse)
library(rgdal)
library(sp)
library(cowplot)
library(raster)
library(knitr)
library(DT)
library(patchwork)
library(flextable)
library(dbscan)

#source('../scripts/Helper_Functions.R')
#load("../data/clean_hier_unit_datasets.RData")

```

## Habitat Complex `r descr[count,] %>% pull(LCODE)`
#### `r descr[count,] %>% pull(Name)`
`r descr[count,] %>% pull(Description)`

### Photos (if available)

```{r pictures, echo=FALSE, message=FALSE, warning=FALSE}
pictures <- list.files(path = "../images", pattern = unitname, full.names = TRUE) %>% grep(pattern = "XX", value = TRUE)

show_pictures <- length(pictures)>0

```

```{r conditional_pictures, echo=FALSE, results='asis', eval=show_pictures}
for(i in seq_along(pictures)){
  path <-  pictures[i]
  cat(paste0("![Ecosystem photo](", path, ") \n\n"))
}

```


### Distribution

`r descr[count,] %>% pull(Distribution)`

Full + regional maps

```{r full map, fig.width = 12, fig.height = 12, echo=FALSE, message=FALSE, warning=FALSE}

 # Full map
T1 <-  basemap +
    geom_tile(data=unit %>% slice_sample(n = 20000), aes(x=x, y=y, col= LCODE, fill = LCODE), lwd = 0.4) +
    scale_fill_manual(values = c("#E0002D")) + scale_color_manual(values = c("#E0002D")) +
    labs(x = "x", y = "y", col = "habitat", fill = "habitat") + 
    theme(legend.position = c(0.1, 0.1), axis.text = element_blank(), axis.ticks = element_blank())

 units <- purrr::map(acbr_ext, ~df_crop(unit, .x)) %>% 
  `[`(purrr::map_int(., nrow) %>% `>`(5000) %>% which()) %>% 
   purrr::map(~if(length(c(far_outliers(.x$x), far_outliers(.x$y))) > 0){
     return(.x %>% slice(-c(far_outliers(x), far_outliers(y))))
   } else {return(.x)})
 
 if(length(units)== 0){
   units <- purrr::map(acbr_ext, ~df_crop(unit, .x)) %>% 
  `[`(purrr::map_int(., nrow) %>% `>`(100) %>% which()) %>% 
   purrr::map(~if(length(c(far_outliers(.x$x), far_outliers(.x$y))) > 0){
     return(.x %>% slice(-c(far_outliers(x), far_outliers(y))))
   } else {return(.x)})
 }
 
  df.grids <- units %>% lapply(geo_bounds, buff = 0.1)
  
T1 <- T1 + geom_rect(data = data.frame(reduce(df.grids, rbind)), aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), fill = NA, color = "black", linewidth = 0.8, inherit.aes = FALSE)

insets <- purrr::map(df.grids, ~T1  + 
                       coord_sf(xlim = c(.x["xmin"], .x["xmax"]), 
                                ylim = c(.x["ymin"], .x["ymax"]), 
                                expand = FALSE))
  
map_mosaic2(df.grids)


```


### Environment

The Habitat Complex `r unitname` is part of the Major Environmental Unit `r groupname`.

```{r calculate abiotic means, message=FALSE, warning=FALSE, include=FALSE}
ecodatAb <- ecodatAb %>% mutate(f = ifelse(LCODE == unitname, "Habitat", 
                                           ifelse(grepl(pattern = groupname, LCODE), "Maj. Env. Unit", 
                                                  "continent")))
ecodatAb$f <- factor(ecodatAb$f, levels = c("continent", "Maj. Env. Unit", "Habitat"))

abioticMeans <- ecodatAb %>% group_by(name, f) %>% 
  summarise(mean = mean(value, na.rm = T)) %>% 
  pivot_wider(names_from = "f", values_from = "mean") %>% 
  mutate(envgr_cont = ifelse(name == "meanTemp", continent/`Maj. Env. Unit`, `Maj. Env. Unit`/ continent), 
         unit_envgr = ifelse(name == "meanTemp", `Maj. Env. Unit`/Habitat, Habitat / `Maj. Env. Unit`))  

```

Major Environmental Unit `r groupname` is, on average, substantially higher in `r df2txt(abioticMeans, 1.25, type = "a")` than continental Antarctica. It is substantially lower in `r df2txt(abioticMeans, 0.75, type = "a")` than the rest of the continent.

The elevation of `r unitname` ranges from `r round(elev_table$min[count])` to `r round(elev_table$max[count])` metres above sea level, but 90% of its pixels fall above `r round(elev_table$lower_90[count])` and below `r round(elev_table$upper_90[count])` metres. Its median elevation is `r round(elev_table$median[count])` metres.

The Habitat Complex is higher in `r df2txt(abioticMeans, 1.25, type = "b")` and lower in `r df2txt(abioticMeans, 0.75, type = "b")` than the rest of Major Environmental Unit `r groupname`.

#### Distinctiveness of the Habitat Complex from its group and the rest of Antarctica

```{r abiotic boxplots, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=11}
ecodatAb %>% slice(sample(1:nrow(ecodatAb), 40000000)) %>% 
  ggplot(aes(y = value, x = f, fill = f)) + 
  geom_boxplot(outlier.size = 0.2) + facet_wrap(.~name, scales = "free_y") + labs(x = "Category", fill = "Category")
```


### Biota

Most widely sampled species in `r groupname` (found in most pixels)
```{r widespread species, echo=FALSE, message=FALSE, warning=FALSE}
species_sorted <- rownames(PA_distinct)[PA_distinct %>% pull(as.character(typv6[count])) %>% order(decreasing = TRUE)]
counts <- PA_distinct %>% pull(as.character(typv6[count])) %>% sort(decreasing = TRUE)

j <- 0
n <- 0
while(n < 10){
  j <- j + 1
  n <- n + length(which(counts == unique(counts)[j]))
}

most_widespread_species <- sppDat[match(species_sorted[1:n], sppDat$scientific),] %>% dplyr::select(scientific, Functional_group, phylum) %>% data.frame(restricted = species_sorted[1:n] %in% restricted_spp, count = counts[1:n], relative_pct = round(100*counts[1:n]/sum(counts), 2)) %>% filter(relative_pct > 0) %>% 
  setNames(c("scientific name", "functional group", "phylum", "Antarctic specialist", "count", "percent of samp"))

flextable(most_widespread_species) %>%
  width(width = c(1.2, 2.4, 1, 1, 0.6, 1)) %>% 
  fontsize(size = 10, part = "all") %>% font(fontname = "Cambria", part = "all") %>%
  italic(j = 1) %>% theme_zebra()  %>% vline(j=1)
```


```{r calculate biotic means, message=FALSE, warning=FALSE, include=FALSE}

ecodatBi <- ecodatBi %>% mutate(f = ifelse(LCODE == unitname, "Habitat", 
                                           ifelse(grepl(pattern = groupname, LCODE), "Maj. Env. Unit", 
                                                  "continent")))
ecodatBi$f <- factor(ecodatBi$f, levels = c("continent", "Maj. Env. Unit", "Habitat"))


bioticMeans <- ecodatBi %>% group_by(taxon, f) %>% 
  summarise(mean = mean(value, na.rm = T)) %>% 
  pivot_wider(names_from = "f", values_from = "mean") %>% 
  mutate(envgr_cont = continent/`Maj. Env. Unit`, 
         unit_envgr = `Maj. Env. Unit`/Habitat )

```  

Major Environmental Unit `r groupname` is, on average, substantially higher in suitability for 
`r df2txt(bioticMeans, 1.25, "taxon", "a")` 
functional groups than continental Antarctica. It is substantially lower in suitability for
`r df2txt(bioticMeans, 0.75, "taxon", "a")` 
than the rest of the continent.

Habitat Complex `r unitname` is higher in suitability for `r df2txt(bioticMeans, 1.25, "taxon", "b")` 
and lower in suitability for `r df2txt(bioticMeans, 0.75, "taxon", "b")` than the rest of Major Environmental Unit `r groupname`.



#### Distinctiveness of the unit from the Major Environmental Unit and the rest of Antarctica 
 
```{r echo=FALSE, fig.height=9, fig.width=10, message=FALSE, warning=FALSE}
ecodatBi %>% slice(sample(1:nrow(ecodatBi), 50000000)) %>% 
  ggplot(aes(y = value, x = f, fill = f)) + 
  geom_boxplot(outlier.size = 0.2) + facet_wrap(.~taxon, scales = "free") + labs(x = "", fill = "") + theme(strip.text = element_text(size = 8), axis.text.x = element_blank())

```
